---
title: "About"
---

## About Building Spatial Datasets starters

This is a [Quarto](https://quarto.org) project repository of starter templates for the [GES668: Building Spatial Datasets](https://bldgspatialdata.github.io/website).

Additional starters I plan to add include:

- Making interactive maps with `{leaflet}` and `{mapview}`
- Analyzing Census data with `{tidycensus}` and `{gt}`

## Presentation starters

### Final Project Presentation

Run the following code in your course repository to download the [final presentation template](//final-project-presentation.qmd) to your current working directory:

```{r}
#| eval: false
usethis::use_github_file("bldgspatialdata/starters", "final-project-presentation.qmd")
```

## Map starters

### Making interactive maps with `{rdeck}`

Take [a look at the template](//rdeck-interactive-map.qmd) or download the starter using the following code:

```{r}
#| eval: false
usethis::use_github_file("bldgspatialdata/starters", "rdeck-interactive-map.qmd")
```
