---
title: "About"
---

## About Building Spatial Datasets starters

This project is a repository of starter templates for students in the UMBC course [GES668: Building Spatial Datasets](https://bldgspatialdata.github.io/website).

Here are a few more starters I plan to add:

-   Making interactive maps with `{leaflet}` and `{mapview}`
-   Analyzing Census data with `{tidycensus}` and `{gt}`
-   Writing a README file for a dataset, analysis, or package

Suggest ideas for other starters or templates by opening an issue in the [project GitHub repository](https://github.com/bldgspatialdata/starters).

## Presentation starters

### Final Project Presentation

Run the following code in your course repository to download the [final presentation template](//final-project-presentation.qmd) to your current working directory:

```{r}
#| eval: false
usethis::use_github_file("bldgspatialdata/starters", "final-project-presentation.qmd")
```

## Map starters

### Making interactive maps with `{rdeck}`

Take [a look at the template](//rdeck-interactive-map.qmd) or download the starter using the following code:

```{r}
#| eval: false
usethis::use_github_file("bldgspatialdata/starters", "rdeck-interactive-map.qmd")
```
