---
title: "Interactive Maps with `{rdeck}`"
format:
  html:
    code-fold: true
execute: 
  echo: true
---

Run the following code in your course or project repository to download the {{< meta title >}} starter to your current working directory:

```{r}
#| eval: false
#| code-fold: false
usethis::use_github_file("bldgspatialdatasets/starters", "rdeck-interactive-map.qmd")
```

## Setup

Using this template requires the `{rdeck}` package which can be installed from GitHub:

```{r}
#| eval: false
#| code-fold: false
pak::install_github("qfes/rdeck@*release")
```

:::{.callout-important appearance='simple' icon=true}
## Sign up for a Mapbox account and get a token
A [Mapbox account](https://account.mapbox.com/auth/signup) and [mapbox access token](https://docs.mapbox.com/help/glossary/access-token) is required for Mapbox basemaps, with or without the Mapbox data service. See the `{rdeck}` documentation on [`mapbox_access_token`](https://qfes.github.io/rdeck/reference/mapbox_access_token.html) for more information.
:::

The template also uses the `{tidycensus}` and `{sf}` packages.

```{r}
#| label: setup
#| code-fold: false
#| output: false
library(rdeck)
library(sf)
library(tidycensus)

options(tigris_use_cache = TRUE)
```

## Map

This template includes code cells downloading and preparing the data, a code cell for creating a rdeck map instance, and two example map code cell you can adapt. The initial code cell also uses the `output: false` to hide messages from the `get_acs()` function.

```{r}
#| code-summary: Download ACS data and prepare data for mapping
#| output: false

# Download data on population by Maryland county
md_population <- get_acs("county",
  state = "MD",
  table = "B01003",
  geometry = TRUE
)

# Transform the data into a geodetic CRS (required by all rdeck layers)
md_population <- st_transform(md_population, 4326)

# Rename the geometry column (required by add_polygon_layer())
md_population <- st_set_geometry(md_population, "polygon")

# Download data on race/ethnic group population by Maryland county
md_race <- get_acs("county",
  state = "MD",
  variables = c(
    "White" = "B03002_003",
    "Black" = "B03002_004",
    "Latino" = "B03002_012"
  ),
  geometry = TRUE
)

# Convert data into a dot density format
md_race_dot_density <- md_race |>
  as_dot_density(
    value = "estimate",
    values_per_dot = 100,
    group = "variable"
  )

# Transform the data into a geodetic CRS and rename the geometry column (required by all rdeck layers)
md_race_dot_density <- st_transform(md_race_dot_density, 4326)

# Rename the geometry column (required by add_scatterplot_layer())
md_race_dot_density <- st_set_geometry(md_race_dot_density, "position")
```

```{r}
#| code-summary: Create light and dark rdeck map instances

# Create a rdeck map instance using the Mapbox light and Mapbox dark themes
md_rdeck_light <- rdeck(
  map_style = mapbox_light(),
  theme = "kepler",
  initial_bounds = st_bbox(md_population),
  height = 600
)

md_rdeck_dark <- rdeck(
  map_style = mapbox_dark(),
  theme = "kepler",
  initial_bounds = st_bbox(md_population),
  height = 600
)
```

Both map code cells use the `column: screen-inset-shaded` option to make the map full width in an HTML output format. Learn more about how to modify the [article layout in a Quarto document](https://quarto.org/docs/authoring/article-layout.html).

This first map uses the `rdeck::add_polygon_layer()` and `rdeck::scale_color_linear()`. Setting `pickable = TRUE` allows you to click on a polygon and see the fields supplied to the `tooltip` parameter.

```{r}
#| code-summary: Create a polygon layer map
#| column: screen-inset-shaded
md_rdeck_light |>
  add_polygon_layer(
    name = "Maryland counties",
    data = md_population,
    opacity = 0.6,
    get_fill_color = scale_color_linear(
      estimate,
      col_label = "Population"
      ),
    tooltip = c(NAME, GEOID, estimate),
    pickable = TRUE
  )
```

This second map uses the `rdeck::add_scatterplot_layer()` and `rdeck::scale_color_category()`. It also uses a custom palette for qualitative data created with the `scales::brewer_pal()` function from the `{scales}` package.
 
```{r}
#| code-summary: Create a scatterplot layer map
#| column: screen-inset-shaded
md_rdeck_dark |>
  add_scatterplot_layer(
    name = "Maryland residents (100 per dot)",
    data = md_race_dot_density,
    radius_min_pixels = 0.75,
    opacity = 0.6,
    get_fill_color = scale_color_category(
      variable,
      palette = scales::brewer_pal("qual"),
      col_label = "Race/ethnicity"
      )
  )
```
